---
title: "Relationship between MBTI personality types and physical traits"
author: "Tina Bakić, Luka Korov, Dora Baričević, Roko Grbelja"
date: "Dec 13, 2023"
output:
  html_document:
    theme: darkly
    toc: yes
    toc_float:
      collapsed: true
subtitle: ""
---

# Introduction

## Libraries

```{r}
library(hexbin)
```

## Dataset

Dataset contains physical and personality data of 97 randomly sampled subjects. All subjects were French-Canadian, living in Canada between the Québec and Sorel-Tracy areas. The Biotonix analyses and report were used on the subjects being tested in order to distinguish postural deviations. Personality was determined by using the Myers-Briggs Type Indicator questionnaire.

Gathered data includes following characteristics for all subjects:

```{r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "

|Column name    |Description                    |
|---------------|-------------------------------|
|AGE            |age in years|
|HEIGHT         |height in inches|
|WEIGHT         |weight in pounds|
|SEX            |biological gender (female/male)|
|ACTIVITY.LEVEL |physical activity level (low/moderate/high)|
|PAIN.1         |level of neck pain (0-100)|
|PAIN.2         |level of pain in upper back (0-100)|
|PAIN.3         |level of pain in middle back (0-100)|
|PAIN.4         |level of pain in lower back (0-100)|
|MBTI           |Myers-Briggs personality type|
|E              |level of extraversion|
|I              |level of introversion|      
|S              |level of sensing|
|N              |level of intuition|
|T              |level of thinking|
|F              |level of feeling|
|J              |level of judging|
|P              |level of perceiving|
|POSTURE        |posutre type (A - ideal, B - kyphosis-lordosis, C - flat back, D - sway-back)|
"
cat(tabl)
```

```{r}
dataset_mbti = read.csv("datasets/MBTI.csv")
```

### Transformations

1. to metric system + BMI

```{r}
dataset_mbti$HEIGHT_CM <- dataset_mbti$HEIGHT * 2.54
dataset_mbti$WEIGHT_KG <- dataset_mbti$WEIGHT * 0.453592
dataset_mbti$BMI <- dataset_mbti$WEIGHT_KG / (dataset_mbti$HEIGHT_CM/100)^2
```

2. separating each personality trait

```{r}
dataset_mbti$MBTI_EI <- substring(dataset_mbti$MBTI, 1, 1)
dataset_mbti$MBTI_SN <- substring(dataset_mbti$MBTI, 2, 2)
dataset_mbti$MBTI_TF <- substring(dataset_mbti$MBTI, 3, 3)
dataset_mbti$MBTI_JP <- substring(dataset_mbti$MBTI, 4, 4)
```

3. from categories to numerical

```{r}
dataset_mbti$SEX_INT <- match(dataset_mbti$SEX, c("Female", "Male"))
dataset_mbti$ACTIVITY.LEVEL_INT <- match(dataset_mbti$ACTIVITY.LEVEL, c("Low", "Moderate", "High"))
dataset_mbti$POSTURE_INT <- match(dataset_mbti$POSTURE, c("A", "B", "C", "D"))
```

```{r}
plot(dataset_mbti$E, dataset_mbti$I)
plot(dataset_mbti$S, dataset_mbti$N)
plot(dataset_mbti$T, dataset_mbti$F)
plot(dataset_mbti$J, dataset_mbti$P)
```

## Research questions

1. Is there a connection between personality type and posture?
2. Are physically attractive people also more extraverted?
3. Is there a difference in height/weight based on personality type?
4. Can you predict a personality type or a specific personality characteristic based on physical traits?

5. Are women more feeling than men?
6. Is there a connection between generation (age) and personality types?


# RQ-1: Is there a connection between personality type and posture?

*[HR]* Postoji li veza izmedu tipa ličnosti i načina držanja?

```{r}
table(
  dataset_mbti$POSTURE,
  dataset_mbti$MBTI
  )

table(
  dataset_mbti$POSTURE,
  dataset_mbti$MBTI_EI
  )


table(
  dataset_mbti$POSTURE,
  dataset_mbti$MBTI_SN
  )

table(
  dataset_mbti$POSTURE,
  dataset_mbti$MBTI_TF
  )


table(
  dataset_mbti$POSTURE,
  dataset_mbti$MBTI_JP
  )
```

## Conclusion

There is a significant ...


# RQ-2: Are physically attractive people also more extraverted?

*[HR]* Jesu li fizički aktivniji ljudi također i ekstrovertiraniji?

Define physically attractive

Conduct independency test


## Conclusion
Physically attractive people are/aren't also more extraverted.


# RQ-3: Is there a difference in height/weight based on personality type?

*[HR]* Postoji li razlika u visini/težini ljudi s obzirom na tip ličnosti?

Box plot check

```{r}
boxplot(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$HEIGHT_CM, 
        dataset_mbti[dataset_mbti$MBTI_EI == "I",]$HEIGHT_CM,
        names = c("Extraverts", "Introverts"),
        main = "Height for extra/introversion")

boxplot(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$WEIGHT_KG, 
        dataset_mbti[dataset_mbti$MBTI_EI == "I",]$WEIGHT_KG,
        names = c("Extraverts", "Introverts"),
        main = "Weight for extra/introversion")

boxplot(dataset_mbti[dataset_mbti$MBTI_SN == "S",]$HEIGHT_CM, 
        dataset_mbti[dataset_mbti$MBTI_SN == "N",]$HEIGHT_CM,
        names = c("Sensing", "Intuition"),
        main = "Height for sensing/intuition")

boxplot(dataset_mbti[dataset_mbti$MBTI_SN == "S",]$WEIGHT_KG, 
        dataset_mbti[dataset_mbti$MBTI_SN == "N",]$WEIGHT_KG,
        names = c("Sensing", "Intuition"),
        main = "Weight for sensing/intuition")
```

Normality check

```{r}
qqnorm(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$WEIGHT_KG, 
       pch = 1, frame = FALSE,
       main='Weight of extraverts')
qqline(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$WEIGHT_KG, 
       col = "steelblue", lwd = 2)

qqnorm(dataset_mbti[dataset_mbti$MBTI_EI == "I",]$WEIGHT_KG, 
       pch = 1, frame = FALSE,
       main='Weight of introverts')
qqline(dataset_mbti[dataset_mbti$MBTI_EI == "I",]$WEIGHT_KG, 
       col = "steelblue", lwd = 2)
```

Variant equality check

```{r}
var.test(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$WEIGHT_KG, dataset_mbti[dataset_mbti$MBTI_EI == "I",]$WEIGHT_KG)
```

T-test with equal/different variants

```{r}
t.test(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$WEIGHT_KG, dataset_mbti[dataset_mbti$MBTI_EI == "I",]$WEIGHT_KG, alt = "greater", var.equal = TRUE)
```

## Conclusion


# RQ-4: Can you predict a personality type or a specific personality characteristic based on physical traits?

*[HR]* Možete li pomoću danih značajki naslutiti tip ličnosti ili rezultat na pojedinoj karakteristici?

Linear regression for E/I values

```{r}
plot(dataset_mbti$AGE, dataset_mbti$E)
plot(dataset_mbti$HEIGHT_CM, dataset_mbti$E)
plot(dataset_mbti$WEIGHT_KG, dataset_mbti$E)
plot(dataset_mbti$BMI, dataset_mbti$E)
boxplot(dataset_mbti[dataset_mbti$SEX == "Female",]$E, 
        dataset_mbti[dataset_mbti$SEX == "Male",]$E,
        names = c("Female", "Male"),
        main = "Extraversion per gender")
boxplot(dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Low",]$E, 
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Moderate",]$E,
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "High",]$E,
        names = c("Low", "Moderate", "High"),
        main = "Extraversion per activity level")
boxplot(dataset_mbti[dataset_mbti$POSTURE == "A",]$E, 
        dataset_mbti[dataset_mbti$POSTURE == "B",]$E, 
        dataset_mbti[dataset_mbti$POSTURE == "C",]$E, 
        dataset_mbti[dataset_mbti$POSTURE == "D",]$E,
        names = c("A", "B", "C", "D"),
        main = "Extraversion per posture")

plot(dataset_mbti$AGE, dataset_mbti$J)
plot(dataset_mbti$HEIGHT_CM, dataset_mbti$J)
plot(dataset_mbti$WEIGHT_KG, dataset_mbti$J)
plot(dataset_mbti$BMI, dataset_mbti$J)
boxplot(dataset_mbti[dataset_mbti$SEX == "Female",]$J, 
        dataset_mbti[dataset_mbti$SEX == "Male",]$J,
        names = c("Female", "Male"),
        main = "Judging per gender")
boxplot(dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Low",]$J, 
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Moderate",]$J,
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "High",]$J,
        names = c("Low", "Moderate", "High"),
        main = "Judging per activity level")
boxplot(dataset_mbti[dataset_mbti$POSTURE == "A",]$J, 
        dataset_mbti[dataset_mbti$POSTURE == "B",]$J, 
        dataset_mbti[dataset_mbti$POSTURE == "C",]$J, 
        dataset_mbti[dataset_mbti$POSTURE == "D",]$J,
        names = c("A", "B", "C", "D"),
        main = "Judging per posture")


plot(dataset_mbti$AGE, dataset_mbti$T)
plot(dataset_mbti$HEIGHT_CM, dataset_mbti$T)
plot(dataset_mbti$WEIGHT_KG, dataset_mbti$T)
plot(dataset_mbti$BMI, dataset_mbti$T)
boxplot(dataset_mbti[dataset_mbti$SEX == "Female",]$T, 
        dataset_mbti[dataset_mbti$SEX == "Male",]$T,
        names = c("Female", "Male"),
        main = "Judging per gender")
boxplot(dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Low",]$T, 
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Moderate",]$T,
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "High",]$T,
        names = c("Low", "Moderate", "High"),
        main = "Judging per activity level")
boxplot(dataset_mbti[dataset_mbti$POSTURE == "A",]$T, 
        dataset_mbti[dataset_mbti$POSTURE == "B",]$T, 
        dataset_mbti[dataset_mbti$POSTURE == "C",]$T, 
        dataset_mbti[dataset_mbti$POSTURE == "D",]$T,
        names = c("A", "B", "C", "D"),
        main = "Judging per posture")
```

Use fastDummies for posture, beware of co-linear columns

Based on regression coefficients, calculate the expected value for certain physical trait

```{r}
summary(lm(E ~ POSTURE_INT + PAIN.1 + PAIN.4, dataset_mbti))

summary(lm(J ~ POSTURE_INT, dataset_mbti))

summary(lm(POSTURE_INT ~ E + S + T + J, dataset_mbti))
summary(lm(ACTIVITY.LEVEL_INT ~ E + S + T + J, dataset_mbti))
summary(lm(SEX_INT ~ E + S + T + J, dataset_mbti))
```

Logistic regression for categories

```{r}

```


## Conclusion

Based on posture/height/weight/something we can predict a personality characteristic...


# Discussion



# Resources

1. Sylvain Guimond, Wael Massrieh: Intricate Correlation between Body Posture, Personality Trait and Incidence of Body Pain: A Cross-Referential Study Report (https://doi.org/10.1371/journal.pone.0037450)
