---
title: "Relationship between MBTI personality types and physical traits"
author: "Tina Bakić, Luka Korov, Dora Baričević, Roko Grbelja"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: darkly
    toc: yes
    toc_float:
      collapsed: true
  pdf_document:
    toc: true
    number_sections: true
subtitle: "Project task for Statistical Data Analysis course at University of Zagreb, Faculty of Electrical Engineering and Computing"
---

# Introduction

## Libraries

```{r message=FALSE, warnings=FALSE}
library(hexbin)
library(dplyr)
```

## Dataset

Dataset contains physical and personality data of 97 randomly sampled subjects. All subjects were French-Canadian, living in Canada between the Québec and Sorel-Tracy areas. The Biotonix analyses and report were used on the subjects being tested in order to distinguish postural deviations. Personality was determined by using the Myers-Briggs Type Indicator questionnaire.

Gathered data includes following characteristics for all subjects:

```{r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "

|Column name    |Description                    |
|---------------|-------------------------------|
|AGE            |age in years|
|HEIGHT         |height in inches|
|WEIGHT         |weight in pounds|
|SEX            |biological gender (female/male)|
|ACTIVITY.LEVEL |physical activity level (low/moderate/high)|
|PAIN.1         |level of neck pain (0-100)|
|PAIN.2         |level of pain in upper back (0-100)|
|PAIN.3         |level of pain in middle back (0-100)|
|PAIN.4         |level of pain in lower back (0-100)|
|MBTI           |Myers-Briggs personality type|
|E              |level of extraversion|
|I              |level of introversion|      
|S              |level of sensing|
|N              |level of intuition|
|T              |level of thinking|
|F              |level of feeling|
|J              |level of judging|
|P              |level of perceiving|
|POSTURE        |posutre type (A - ideal, B - kyphosis-lordosis, C - flat back, D - sway-back)|
"
cat(tabl)
```

```{r}
dataset_mbti <- read.csv("datasets/MBTI.csv")
```

### Transformations

1.  to metric system + BMI

```{r}
dataset_mbti$HEIGHT_CM <- dataset_mbti$HEIGHT * 2.54
dataset_mbti$WEIGHT_KG <- dataset_mbti$WEIGHT * 0.453592
dataset_mbti$BMI <- dataset_mbti$WEIGHT_KG / (dataset_mbti$HEIGHT_CM / 100)^2
```

2.  separating each personality trait

```{r}
dataset_mbti$MBTI_EI <- substring(dataset_mbti$MBTI, 1, 1)
dataset_mbti$MBTI_SN <- substring(dataset_mbti$MBTI, 2, 2)
dataset_mbti$MBTI_TF <- substring(dataset_mbti$MBTI, 3, 3)
dataset_mbti$MBTI_JP <- substring(dataset_mbti$MBTI, 4, 4)
```

3.  from categories to numerical

```{r}
dataset_mbti$SEX_INT <- match(dataset_mbti$SEX, c("Female", "Male"))
dataset_mbti$ACTIVITY.LEVEL_INT <- match(dataset_mbti$ACTIVITY.LEVEL, c("Low", "Moderate", "High"))
dataset_mbti$POSTURE_INT <- match(dataset_mbti$POSTURE, c("A", "B", "C", "D"))
```

4.  grouping similar personality types

```{r}
dataset_mbti %>%
        mutate(MBTI_GROUP = case_when(
                MBTI == "INTJ" | MBTI == "INTP" | MBTI == "ENTJ" | MBTI == "ENTP" ~ "Analyst",
                MBTI == "INFJ" | MBTI == "INFP" | MBTI == "ENFJ" | MBTI == "ENFP" ~ "Diplomat",
                MBTI == "ISTJ" | MBTI == "ISFJ" | MBTI == "ESTJ" | MBTI == "ESFJ" ~ "Sentinel",
                MBTI == "ISTP" | MBTI == "ISFP" | MBTI == "ESTP" | MBTI == "ESFP" ~ "Explorer"
        ))
```

```{r}
plot(dataset_mbti$E, dataset_mbti$I)
plot(dataset_mbti$S, dataset_mbti$N)
plot(dataset_mbti$T, dataset_mbti$F)
plot(dataset_mbti$J, dataset_mbti$P)
```

## Research questions

1.  Is there a connection between personality type and posture?

2.  Are physically active people also more extroverted?

3.  Is there a difference in height/weight based on personality type?

4.  Can you predict a personality type or a specific personality characteristic based on physical traits?

5.  Are women more feeling than men?

# RQ-1: Is there a connection between personality type and posture?

*[HR]* Postoji li veza izmedu tipa ličnosti i načina držanja?

## For MBTI Personality Type:

$H_0:$ There is no association between the individual personality trait and posture.

$H_1:$ There is an association between the individual personality trait and posture.

```{r}
mbti_posture_table <- table(dataset_mbti$POSTURE, dataset_mbti$MBTI)
mbti_posture_table

chi_squared_test_mbti <- chisq.test(mbti_posture_table)
chi_squared_test_mbti
```

$p = 0.001072 < \alpha = 0.05$

### Conclusion

The p-value of the Chi-squared test is lower than the significance level, therefore, we reject the null hypothesis and conclude that there is a significant correlation between an individual's personality type and their posture. It is important to note that the Chi-squared approximation may not be correct as our sample size is relatively small and our contingency table is large and data is sparse. We can try and remedy this by aggregating the posture types into two categories: GOOD and BAD.

```{r}
dataset_mbti$POSTURE_AGG <- ifelse(dataset_mbti$POSTURE == "A", "GOOD", "BAD")
table(dataset_mbti$POSTURE_AGG)
```

```{r}
mbti_posture_agg_table <- table(dataset_mbti$POSTURE_AGG, dataset_mbti$MBTI)

chi_squared_test_mbti_agg <- chisq.test(mbti_posture_agg_table)

chi_squared_test_mbti_agg

```

## For individual personality traits:

$H_0:$ There is no association between the individual personality trait and posture.

$H_1:$ There is an association between the individual personality trait and posture.

```{r}
ei_posture_table <- table(dataset_mbti$POSTURE, dataset_mbti$MBTI_EI)
sn_posture_table <- table(dataset_mbti$POSTURE, dataset_mbti$MBTI_SN)
tf_posture_table <- table(dataset_mbti$POSTURE, dataset_mbti$MBTI_TF)
jp_posture_table <- table(dataset_mbti$POSTURE, dataset_mbti$MBTI_JP)
```

```{r}
chi_squared_test_ei <- chisq.test(ei_posture_table)
chi_squared_test_ei
```

```{r}
chi_squared_test_sn <- chisq.test(sn_posture_table)
chi_squared_test_sn
```

```{r}
chi_squared_test_tf <- chisq.test(tf_posture_table)
chi_squared_test_tf
```

```{r}
chi_squared_test_jp <- chisq.test(jp_posture_table)
chi_squared_test_jp
```

### Conclusion

Out of the 4 personality traits, only the extroversion/introversion trait passes the chi-squared test on the significance level of $0.05$. We reject the null hypothesis for the extroversion/introversion personality trait and conclude that there is a significant connection between an individual's extrovertedness and their posture.

If we try and visualize the extroversion scores for each posture type, it is apparent that extroverted individuals tend to have better posture.

```{r}
boxplot(dataset_mbti[dataset_mbti$POSTURE == "A", ]$E,
        dataset_mbti[dataset_mbti$POSTURE == "B", ]$E,
        dataset_mbti[dataset_mbti$POSTURE == "C", ]$E,
        dataset_mbti[dataset_mbti$POSTURE == "D", ]$E,
        names = c("Ideal", "Kyphosis-lordosis", "Flat back", "Sway-back"),
        main = "Extroversion per posture type"
)
```

# RQ-2: Are physically active people also more extroverted?

*[HR]* Jesu li fizički aktivniji ljudi također i ekstrovertiraniji?

```{r}
boxplot(dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Low",]$E, 
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Moderate",]$E,
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "High",]$E,
        names = c("Low", "Moderate", "High"),
        main = "Extraversion per activity level")
```

## ANOVA

$H_0:$ There is no significant difference in mean extraversion scores among people with different activity levels (Low, Moderate, High).

$H_1:$ At least one pair of activity levels has significantly different mean extraversion scores.

```{r}
extraversion_scores <- dataset_mbti$E[dataset_mbti$MBTI_EI == "E"]

low_activity <- dataset_mbti$E[dataset_mbti$ACTIVITY.LEVEL == "Low"]
moderate_activity <- dataset_mbti$E[dataset_mbti$ACTIVITY.LEVEL == "Moderate"]
high_activity <- dataset_mbti$E[dataset_mbti$ACTIVITY.LEVEL == "High"]

anova_result <- aov(E ~ ACTIVITY.LEVEL, data = dataset_mbti)

summary(anova_result)
```

$p = 0.555 > \alpha = 0.05$

We cannot reject null hypothesis. There isn't enough evidence to conclude that there are significant differences in extraversion scores among individuals with different activity levels.

## Independency

$H_0:$ Activity levels and extraversion are independent of each other.

$H_1:$ Activity levels and extraversion are not independent of each other, there is a relationship between them.

```{r}
contingency_table <- table(dataset_mbti$ACTIVITY.LEVEL, dataset_mbti$MBTI_EI)
contingency_table
```

```{r}
chi_sq_test <- chisq.test(contingency_table)
print(chi_sq_test)
```

$p = 0.6562 > \alpha = 0.05$

We cannot reject null hypothesis. There isn't enough evidence to conclude that extraversion and different activity levels are not independant.

## Conclusion

There is not enough evidence to say that physically active people are also more extroverted.

# RQ-3: Is there a difference in height/weight based on personality type?

*[HR]* Postoji li razlika u visini/težini ljudi s obzirom na tip ličnosti?

Box plot check

```{r}
boxplot(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$HEIGHT_CM, 
        dataset_mbti[dataset_mbti$MBTI_EI == "I",]$HEIGHT_CM,
        names = c("Extraverts", "Introverts"),
        main = "Height for extra/introversion")

boxplot(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$WEIGHT_KG, 
        dataset_mbti[dataset_mbti$MBTI_EI == "I",]$WEIGHT_KG,
        names = c("Extraverts", "Introverts"),
        main = "Weight for extra/introversion")

boxplot(dataset_mbti[dataset_mbti$MBTI_SN == "S",]$HEIGHT_CM, 
        dataset_mbti[dataset_mbti$MBTI_SN == "N",]$HEIGHT_CM,
        names = c("Sensing", "Intuition"),
        main = "Height for sensing/intuition")

boxplot(dataset_mbti[dataset_mbti$MBTI_SN == "S",]$WEIGHT_KG, 
        dataset_mbti[dataset_mbti$MBTI_SN == "N",]$WEIGHT_KG,
        names = c("Sensing", "Intuition"),
        main = "Weight for sensing/intuition")
```

Normality check

```{r}
qqnorm(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$WEIGHT_KG, 
       pch = 1, frame = FALSE,
       main='Weight of extraverts')
qqline(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$WEIGHT_KG, 
       col = "steelblue", lwd = 2)

qqnorm(dataset_mbti[dataset_mbti$MBTI_EI == "I",]$WEIGHT_KG, 
       pch = 1, frame = FALSE,
       main='Weight of introverts')
qqline(dataset_mbti[dataset_mbti$MBTI_EI == "I",]$WEIGHT_KG, 
       col = "steelblue", lwd = 2)
```

Variant equality check

```{r}
var.test(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$WEIGHT_KG, dataset_mbti[dataset_mbti$MBTI_EI == "I",]$WEIGHT_KG)
```

T-test with equal/different variants

```{r}
t.test(dataset_mbti[dataset_mbti$MBTI_EI == "E",]$WEIGHT_KG, dataset_mbti[dataset_mbti$MBTI_EI == "I",]$WEIGHT_KG, alt = "greater", var.equal = TRUE)
```

## Conclusion

# RQ-4: Can you predict a personality type or a specific personality characteristic based on physical traits?

*[HR]* Možete li pomoću danih značajki naslutiti tip ličnosti ili rezultat na pojedinoj karakteristici?

Linear regression for E/I values

```{r}
plot(dataset_mbti$AGE, dataset_mbti$E)
plot(dataset_mbti$HEIGHT_CM, dataset_mbti$E)
plot(dataset_mbti$WEIGHT_KG, dataset_mbti$E)
plot(dataset_mbti$BMI, dataset_mbti$E)
boxplot(dataset_mbti[dataset_mbti$SEX == "Female",]$E, 
        dataset_mbti[dataset_mbti$SEX == "Male",]$E,
        names = c("Female", "Male"),
        main = "Extraversion per gender")
boxplot(dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Low",]$E, 
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Moderate",]$E,
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "High",]$E,
        names = c("Low", "Moderate", "High"),
        main = "Extraversion per activity level")
boxplot(dataset_mbti[dataset_mbti$POSTURE == "A",]$E, 
        dataset_mbti[dataset_mbti$POSTURE == "B",]$E, 
        dataset_mbti[dataset_mbti$POSTURE == "C",]$E, 
        dataset_mbti[dataset_mbti$POSTURE == "D",]$E,
        names = c("A", "B", "C", "D"),
        main = "Extraversion per posture")

plot(dataset_mbti$AGE, dataset_mbti$J)
plot(dataset_mbti$HEIGHT_CM, dataset_mbti$J)
plot(dataset_mbti$WEIGHT_KG, dataset_mbti$J)
plot(dataset_mbti$BMI, dataset_mbti$J)
boxplot(dataset_mbti[dataset_mbti$SEX == "Female",]$J, 
        dataset_mbti[dataset_mbti$SEX == "Male",]$J,
        names = c("Female", "Male"),
        main = "Judging per gender")
boxplot(dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Low",]$J, 
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Moderate",]$J,
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "High",]$J,
        names = c("Low", "Moderate", "High"),
        main = "Judging per activity level")
boxplot(dataset_mbti[dataset_mbti$POSTURE == "A",]$J, 
        dataset_mbti[dataset_mbti$POSTURE == "B",]$J, 
        dataset_mbti[dataset_mbti$POSTURE == "C",]$J, 
        dataset_mbti[dataset_mbti$POSTURE == "D",]$J,
        names = c("A", "B", "C", "D"),
        main = "Judging per posture")


plot(dataset_mbti$AGE, dataset_mbti$T)
plot(dataset_mbti$HEIGHT_CM, dataset_mbti$T)
plot(dataset_mbti$WEIGHT_KG, dataset_mbti$T)
plot(dataset_mbti$BMI, dataset_mbti$T)
boxplot(dataset_mbti[dataset_mbti$SEX == "Female",]$T, 
        dataset_mbti[dataset_mbti$SEX == "Male",]$T,
        names = c("Female", "Male"),
        main = "Judging per gender")
boxplot(dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Low",]$T, 
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "Moderate",]$T,
        dataset_mbti[dataset_mbti$ACTIVITY.LEVEL == "High",]$T,
        names = c("Low", "Moderate", "High"),
        main = "Judging per activity level")
boxplot(dataset_mbti[dataset_mbti$POSTURE == "A",]$T, 
        dataset_mbti[dataset_mbti$POSTURE == "B",]$T, 
        dataset_mbti[dataset_mbti$POSTURE == "C",]$T, 
        dataset_mbti[dataset_mbti$POSTURE == "D",]$T,
        names = c("A", "B", "C", "D"),
        main = "Judging per posture")
```

Use fastDummies for posture and MBTI gruops, beware of co-linear columns

Based on regression coefficients, calculate the expected value for certain physical trait

```{r}
summary(lm(E ~ POSTURE_INT + PAIN.1 + PAIN.4 + HEIGHT_CM + WEIGHT_KG + F, dataset_mbti))

summary(lm(J ~ POSTURE_INT, dataset_mbti))

summary(lm(POSTURE_INT ~ E + S + T + J, dataset_mbti))
summary(lm(ACTIVITY.LEVEL_INT ~ E + S + T + J, dataset_mbti))
summary(lm(SEX_INT ~ E + S + T + J, dataset_mbti))
```

Logistic regression for categories

```{r}

```

## Conclusion

Based on posture/height/weight/something we can predict a personality characteristic...

# RQ-5: Are women more feeling than men?

*[HR]* Jesu li žene osjećajnije nego muškarci?

## Independency

$H_0:$ There is no difference in the distribution of the Feeling variable between men and women.

$H_1:$ There is a difference in the distribution of the Feeling variable between men and women.

```{r}
contingency_table <- table(dataset_mbti$SEX, dataset_mbti$MBTI_TF)
contingency_table
```

```{r}
chi_sq_test <- chisq.test(contingency_table)
print(chi_sq_test)
```

$p = 0.0002972 > \alpha = 0.05$

We can reject null hypothesis. There is a significant difference in the distribution of the Feeling variable between men and women.

## Test of proportions

$H_0:$ The proportion of individuals classified as Feeling is the same for men and women.

$H_1:$ The proportion of individuals classified as Feeling is higher for women compared to men.

```{r}
feeling_counts <- contingency_table["Female", "F"]
total_counts <- rowSums(contingency_table)

proportion_test <- prop.test(feeling_counts, total_counts["Female"])
proportion_test
```

$p = 6.334e-05 < \alpha = 0.05$

We can reject null hypothesis.

The confidence interval, ranging from $0.6523763$ to $0.8927585$, gives us a likely range for the true proportion of individuals labeled as "Feeling" in the entire population. The sample estimate $p$ is $0.7959184$ and it suggests that in our sample, a substantial proportion of females are classified as "Feeling." If there was no difference, the expected proportion would be $0.5$. In simpler terms, this indicates that in our sample a larger number of females are likely to be classified as "Feeling" rather than not.

## Conclusion

Women are more feeling than men.

# Resources

1.  Sylvain Guimond, Wael Massrieh: Intricate Correlation between Body Posture, Personality Trait and Incidence of Body Pain: A Cross-Referential Study Report (<https://doi.org/10.1371/journal.pone.0037450>)
